### 2022-01-12 manuscript/manuscript.Rmd
Update manuscript.Rmd

# Discussion
In 330 influential journals operating across 22 scientific domains we found substantial variation in how PPPR was handled in both policy and practice. A considerable number of journals (37%) did not offer any PPPR options at all, and many imposed restricted length and time-to-submit limits, representing a serious obstacle to the cultivation, documentation, and dissemination of post-publication critique.

There was substantial variation across scientific domains with Clinical Medicine standing out as the domain that offered the most PPPR options, had the most restrictive PPPR options, and published the most PPPR (for concordant evidence, see Altman, 2002; Von Elm et al., 2009; Winker, 2013). Other health-related domains seemed to have a more active PPPR culture than non-health related domains like the physical and social sciences. Many domains, such as mathematics, showed little evidence of any PPPR activity, with few journals offering PPPR and scarce evidence of published PPPR. Our data do not speak to the causal forces that underlie these inter-domain differences, but some potential contributing factors could be cultural (e.g., different attitudes towards scientific criticism and how it should be handled), pragmatic (e.g., differences in methodological standards and research quality, manifesting in differential need for scientific criticism), bureaucratic (e.g., different resources available to support PPPR), or historic (e.g., individuals or events that highlighted the value of PPPR).

PPPR could usually be mapped to one of three main types (letters, commentaries, and web comments), of which letters were most common in policy and practice. Generally, letters were most restrictive, followed by commentaries, then web comments. Typically, letters had to be shorter, submitted more quickly, and contain fewer references relative to commentaries. Usually, web comments had no stated limits, except for a quarter that had length limits. Policies implied that commentaries were more likely to be sent for independent external peer review, with letters more likely to be handled exclusively by the editorial team. Web comments were typically subject to ‘light’ editorial moderation or no review at all. Some journals may offer less restrictive web comments to compensate for other more restrictive PPPR options they offer (Table 2).

The extent to which journal limits on PPPR are reasonable or unreasonable is a somewhat subjective determination and there are likely to be competing interests between what is best for journals and what is best for the advancement of science. Restrictions on PPPR may arise from editorial bias against criticism of papers they have published. Editors may also prefer to facilitate only what they perceive as the most timely and concise debate. However, length restrictions arbitrarily limit the scope of PPPR, particularly if the criticism involves extended analyses or additional data. One can certainly say very little of substance in 175, 200, or 250 words (the most restrictive length limits). Restricting the number of references to 3, 4, or 5 (the most restrictive reference limits) may prevent links to relevant evidential, contextual, or methodological information, undermining an aspect of scholarship that is surely as important to PPPR as it is to regular articles. Finally, imposing time-to-submit limits on PPPR is clearly not justifiable from a scientific perspective because important critiques may arise at any time. Limiting the time allowed to submit PPPR to 2, 3, or 4 weeks (the most restrictive time-to-submit limits) seems especially unjustifiable and poses a serious obstacle to the publication of scientific critique. An earlier study describing strict length and time-to-submit limits imposed on PPPR at six leading medical journals led the author to trenchantly conclude that "In effect, there is a statute of limitations by which authors of articles in these journals are immune to disclosure of methodological weaknesses once some arbitrary (short) period has elapsed, which cannot be right." (Altman, 2002; also see Altman, 2005).

Our exploration of how PPPR is used in practice suggested that letters are far more common than other types of PPPR, perhaps because they are the most frequently available PPPR option and also because, as formal articles, they may impart greater academic credit to their authors than informal web comments (Table 3). We found that half of PPPRs were behind a paywall, sometimes even when the target article was publicly accessible. This reduces access to PPPR for both professional scientists and other readers, like patients, journalists, and policy-makers (Winker, 2015). Most PPPR had conflict of interest statements and half of those statements declared a potential conflict. Conflict of interest statements enable readers to evaluate risk of bias and seem just as important to PPPR as they are to other academic articles (Dunn et al., 2016).

@@

Many of our findings imply that the extant culture of journal-based post-publication critique is suboptimal, though more detailed scrutiny of PPPR dynamics at specific journals will enhance this diagnosis. It is interesting to note that, of the 123 journals that did not offer PPPR options, 83 were members of COPE, an organisation whose guidelines state that "Journals must allow debate post publication either on their site, through letters to the editor, or on an external moderated site, such as PubPeer" (Committee on Publication Ethics, 2019). Further research is needed to explore the extent to which the current state of PPPR is a result of principled decisions or practical obstacles. It is tempting to look outside of the journal system for solutions to facilitate PPPR (Bastian, 2014); however, attempts to establish dedicated PPPR platforms have met with limited success — one major platform, PubMed Commons, was shut down in 2018 due to low usage (NCBI Insights, 2018). In Box 2, we have suggested some policy changes for journals to consider that may facilitate PPPR.

The cultivation, documentation, and dissemination of post-publication critique is an important part of a healthy and self-correcting research literature. Our study reveals considerable variation in how PPPR is handled by journals operating across scientific domains. Clinical medicine had a much more active PPPR culture than other domains; but it's journals also imposed the strictest limits. Overall, PPPR appears to be tightly controlled and restricted by influential academic journals. At many journals, it was not possible to publish PPPR at all, and journals that did offer PPPR options often imposed strict length and time-to-submit restrictions. The PPPR we did identify appeared to have negligible impact on target article authors. These data provide a stratum of empirical evidence upon which to base debates about how scientific critique should be optimally handled. We encourage stakeholders across the academic ecosystem to explore ways to foster a richer culture of post-publication critique.

>>>>

# Discussion
We found substantial variation in how PPPR was handled in both policy and practice at 330 influential journals operating across 22 scientific domains. Publication of PPPR was rare in most domains and a considerable number of journals (37%) did not offer any options for submitting PPPR. Journals that did offer PPPR often imposed restrictive length and time-to-submit limits. Overall, influential journals often represented a serious obstacle to the cultivation, documentation, and dissemination of post-publication critique.

There was substantial variation across scientific domains with journals in Clinical Medicine standing out as offering the most PPPR options and publishing the most PPPR, but also imposed the most restrictive length and time-to-submit limits (for concordant evidence, see Altman, 2002; Von Elm et al., 2009; Winker, 2013). Other health-related domains seemed to have a more active PPPR culture than non-health related domains like the physical sciences and social sciences. Many domains, such as mathematics, showed little evidence of any PPPR activity, with few journals offering PPPR and scarce evidence of published PPPR. Our data do not speak to the causal forces that underlie these inter-domain differences, but some potential contributing factors could be cultural (e.g., different attitudes towards scientific criticism and how it should be handled), pragmatic (e.g., differences in methodological standards and research quality, manifesting in differential need for scientific criticism), bureaucratic (e.g., different resources available to support PPPR), or historic (e.g., individuals or events that highlighted the value of PPPR).

PPPR could usually be mapped to one of three main types (letters, commentaries, and web comments), of which letters were most common in policy and practice. Generally, letters had the most restrictive limits, followed by commentaries, then web comments. Typically, letters had to be shorter, submitted more quickly, and contain fewer references relative to commentaries. Usually, web comments had no stated limits, except for a quarter that had length limits. Policies implied that commentaries were more likely to be sent for independent external peer review, with letters more likely to be handled exclusively by the editorial team. Web comments were typically subject to ‘light’ editorial moderation or no review at all. Some journals may offer less restrictive web comments to compensate for other more restrictive PPPR options they offer (Table 2).

The extent to which journal limits on PPPR are reasonable or unreasonable is a somewhat subjective determination and there are likely to be competing interests between what is best for journals and what is best for the advancement of science. Restrictions on PPPR may arise from editorial bias against criticism of papers they have published. Editors may also prefer to allow only what they perceive as the most timely and concise debate. However, length restrictions arbitrarily limit the scope of PPPR, particularly if the criticism involves extended analyses or additional data. One can certainly say very little of substance in 175, 200, or 250 words (the most restrictive length limits). Restricting the number of references to 3, 4, or 5 (the most restrictive reference limits) may prevent links to relevant evidential, contextual, or methodological information, undermining an aspect of scholarship that is surely as important to PPPR as it is to regular articles. Finally, imposing time-to-submit limits on PPPR is clearly not justifiable from a scientific perspective because important critiques may arise at any time. Limiting the time allowed to submit PPPR to 2, 3, or 4 weeks (the most restrictive time-to-submit limits) seems especially unjustifiable and poses a serious threat to the dissemination of scientific critique. An earlier study describing strict length and time-to-submit limits imposed on PPPR at six leading medical journals led the author to trenchantly conclude that "In effect, there is a statute of limitations by which authors of articles in these journals are immune to disclosure of methodological weaknesses once some arbitrary (short) period has elapsed, which cannot be right." (Altman, 2002; also see Altman, 2005).

Our exploration of how PPPR is used in practice suggested that letters are far more common than other types of PPPR, perhaps because they are the most frequently available PPPR option and also because, as formal articles, they may impart greater academic credit to their authors than informal web comments (Table 3). We found that half of the PPPRs we examined were behind a paywall, sometimes even when the target article was publicly accessible. This reduces access to PPPR for both professional scientists and other readers, like patients, journalists, and policy-makers (Winker, 2015). Most PPPR had conflict of interest statements and half of those statements declared a potential conflict. Conflict of interest statements enable readers to evaluate an important risk of bias and seem just as relevant to PPPR as they are to other academic articles (Dunn et al., 2016).

@@

Many of our findings imply that the extant culture of journal-based post-publication critique is suboptimal, though more detailed scrutiny of PPPR dynamics at specific journals will enhance this diagnosis. It is interesting to note that, of the 123 journals that did not offer any options for submitting PPPR, 83 were members of COPE, an organisation whose guidelines state that "Journals must allow debate post publication either on their site, through letters to the editor, or on an external moderated site, such as PubPeer" (Committee on Publication Ethics, 2019). Further research is needed to explore the extent to which the current state of PPPR is a result of principled editorial decisions or practical obstacles. It is tempting to look outside of the journal system for solutions to facilitate PPPR (Bastian, 2014); however, attempts to establish dedicated PPPR platforms have met with limited success — one major platform, PubMed Commons, was shut down in 2018 due to low usage (NCBI Insights, 2018). In Box 2, we have suggested some policy changes for journals to consider that may facilitate PPPR.

The cultivation, documentation, and dissemination of post-publication critique is an important part of a healthy and self-correcting research literature. Our study reveals considerable variation in how PPPR is handled by journals operating across scientific domains. Clinical Medicine had a much more active PPPR culture than other domains; but it's journals also imposed the strictest limits. Overall, PPPR appears to be tightly controlled and restricted by influential academic journals. At many journals, it was not possible to publish PPPR at all, and journals that did offer PPPR options often imposed strict length and time-to-submit restrictions. The PPPR we did identify appeared to have negligible impact on target article authors. These data provide a stratum of empirical evidence upon which to base debates about how scientific critique should be optimally handled. We encourage stakeholders across the academic ecosystem to explore ways to foster a richer culture of post-publication critique.
